---
title: "Set up your environment to contribute"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{setup_developers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Set up

In this vignette we explain how to set up your environment to be able to contribute to the package. **OmopPostgresConnector** is an open-source package open for contributions, if you want to contribute you will need to be able to run the tests locally. To do so you need a local postgres instance with the following credentials:

```
POSTGRES_USER=omop_postgres_connector
POSTGRES_PASSWORD=omopverse
POSTGRES_DB=omop_test
```

The GitHub actions run on a container where these [credentials](https://github.com/oxford-pharmacoepi/OmopPostgresConnector/blob/main/.github/workflows/postgres.yml#L13-L15) have already been created, you can rely on the GitHub actions to run tests there.

There are several ways to have this:

### Install Postgres and create manually a database

1. [Download postgres](https://www.postgresql.org/download/)
2. Set up your installer and install Postgre and pgAdmin 4. Please use PORT 5432, if port 5432 is not available create an environment variable `OMOP_POSTGRES_CONNECTOR_PORT` with the port that you are using:
```
OMOP_POSTGRES_CONNECTOR_PORT=5432
```
3. Create your a user. Please use USER: omop_postgres_connector and PASSWORD: omopverse; if this is not possible please create environment variables to provide the credentials:
```
OMOP_POSTGRES_CONNECTOR_USER=omop_postgres_connector
OMOP_POSTGRES_CONNECTOR_PASSWORD=omopverse
```

4. Create a database. Please use DATABASE: omop_test; if this is not possible please create environment variables to provide the credentials:
```
OMOP_POSTGRES_CONNECTOR_DB=omop_test
```

If you successfully completed the 4 steps please go to [Check the set up](add link).

### Install Postgres via docker

1. [Download docker desktop](https://www.docker.com/products/docker-desktop)

2. Run the following code from the terminal:

```
docker run --name local-postgres \
  -e POSTGRES_USER=omop_postgres_connector \
  -e POSTGRES_PASSWORD=omopverse \
  -e POSTGRES_DB=omop_test \
  -p 5432:5432 \
  -d postgres:16
```

## Check the set up

If your connection is successful you should be able to run the following lines of code without any error:

```{r}
library(DBI)
library(RPostgres)
library(dplyr)
library(dbplyr)

# create connection
con <- dbConnect(
  drv = Postgres(),
  dbname = Sys.getenv("OMOP_POSTGRES_CONNECTOR_DB", "omop_test"),
  host = "localhost",
  port = Sys.getenv("OMOP_POSTGRES_CONNECTOR_PORT", "5432"),
  user = Sys.getenv("OMOP_POSTGRES_CONNECTOR_USER", "omop_postgres_connector"),
  password = Sys.getenv("OMOP_POSTGRES_CONNECTOR_PASSWORD", "omopverse")
)
con

# create new schema
dbExecute(conn = con, statement = "CREATE SCHEMA opc_main;")
dbWriteTable(conn = con, name = Id(schema = "opc_main", table = "test"), value = cars)
tbl(con, I("opc_main.test"))
dbExecute(con, "DROP SCHEMA opc_main CASCADE;")

# create new database
dbExecute(conn = con, statement = "CREATE DATABASE opc_test;")
dbDisconnect(conn = con)

# connect to new database
con <- dbConnect(
  drv = Postgres(),
  dbname = "opc_test",
  host = "localhost",
  port = Sys.getenv("OMOP_POSTGRES_CONNECTOR_PORT", "5432"),
  user = Sys.getenv("OMOP_POSTGRES_CONNECTOR_USER", "omop_postgres_connector"),
  password = Sys.getenv("OMOP_POSTGRES_CONNECTOR_PASSWORD", "omopverse")
)

dbExecute(conn = con, statement = "CREATE SCHEMA opc_main;")
dbWriteTable(conn = con, name = Id(schema = "opc_main", table = "test"), value = cars)
tbl(con, I("opc_main.test"))
dbDisconnect(conn = con)

# delete created database
con <- dbConnect(
  drv = Postgres(),
  dbname = Sys.getenv("OMOP_POSTGRES_CONNECTOR_DB", "omop_test"),
  host = "localhost",
  port = Sys.getenv("OMOP_POSTGRES_CONNECTOR_PORT", "5432"),
  user = Sys.getenv("OMOP_POSTGRES_CONNECTOR_USER", "omop_postgres_connector"),
  password = Sys.getenv("OMOP_POSTGRES_CONNECTOR_PASSWORD", "omopverse")
)
dbExecute(con, "DROP DATABASE opc_test;")
dbDisconnect(conn = con)

```

Now that you are ready try cloning the code locally and run `devtools::check()` if no error arises you are all set up.
